On the Minimum Core Mass for Giant Planet Formation

### 2013-02-06 tex/draft/draft.tex
reshuffling of sections in outline
All methods in section 2, which should gradually build from the general to the specific.  Ana, if you're going to start writing here, just keep it as basic and factual as possible and not too wordy.  I.e. just get the basic equations and a few connecting words to the page.  The flowery language if anywhere belongs elsewhere and later modifications are possible.

\shorttitle{Critical Core Mass at Wide Separations}
\shortauthors{Piso \& Youdin}

\title{Lower Limits on the Core Mass for Giant Planets}
%\title{The Critical Core Mass of Wide-Separation Giant Planets:  Lower Limits from Kelvin-Helmholtz Cooling}
\author{Ana-Maria Piso}
\affil{Harvard Smithsonian Center for Astrophysics}
\author{ Andrew N.\ Youdin}
\affil{JILA, University of Colorado}
%\begin{abstract}
%\end{abstract}

\section{Introduction}
\emgr{Background topics to mention include: core accretion history, giant planets at wide separations and theoretical challenges posed (GI probably to big, core accretion maybe to slow).}

@@

\section{Quasistatic Two Layer Model}\label{sec:model}

\subsection{Model Assumptions}
\emgr{ Define and describe basic length scales; Describe disk conditions: MMSN etc. Describe basic assumptions: 1D atmosphere, smooth matching onto the disk, atmosphere composition etc., perhaps in list format; Describe assumptions specific to our model: quasistatic, $L_{KH} \gg L_{acc}$, two-layer, constant $L$. Explain how our model is intermediary between purely static models with $L=L_{acc}$ and full time-dependent evolutionary models. Could be split into sub-subsections.}

\subsection{Structure Equations}
\emgr{Give in radius coordinates. EOS (polytrope, various assumption for $\nabla_{\mathrm{ad}}$ and $\mu$) and opacity laws (noting that we only consider cool radiative zones with dust opacities). Give boundary conditions (probably not worth a whole subsection by itself)}

%\subsection{Boundary Conditions: Core and Disk}

%\subsection{Convective Zone}
%\emgr{might not need a (sub)section}

%\subsection{Radiative Zone Solutions}
%\emgr{Give analytic solution valid for power law opacity and lack of self-gravity.}
%
%\subsection{Matching at the Radiative-Convective Boundary (RCB)}
%\emgr{Include figure of an example atmospheric profile.}
%
%\subsection{The Gravothermal Catastrophe}
%\emgr{Show and explain double valued solutions, including caveat that upper branch breaks several model assumptions.}

\section{Cooling History}
\emgr{Start with basic idea of quasistatic contraction and $L = -\dot{E}$, then write out cooling equation and explain terms. Explain where it comes from (local cooling + virial equilibrium)}
\subsection{Role of Finite Pressure}
 \emgr{Give more more general result, but defer derivation to Appendix \ref{sec:virial}. Perhaps not worth a subsection by itself.}
%\subsection{Estimating a Critical Core Mass}
%\emgr{Describe how and why we use the time to gravothermal catastrophe as estimate of minimum core mass}

\section{Numerical method}
\emgr{Explain the shooting technique, then how to obtain an evolutionary series from cooling model between subsequent static atmospheres. Show P-T, M-T, M-R etc. plots. Also L-t, L-M plots.}
\subsection{The Gravothermal Catastrophe \emgr{or} {Entropy Minimum} \emgr{or} {Negative Luminosity} \emgr{or} Something else?}
\emgr{Describe that time / luminosity becomes negative when $M_{atm} \sim M_c$.}

@@

\section{Discussion}
\subsection{Kelvin-Helmholtz Contraction vs. Planetesimal Accretion}
\emgr{Compare and contrast our results with Rafikov's, show accretion rates plot; discuss why our fixed core mass assumptions are valid in the regime we are considering.}
\subsection{Analytic Cooling Model}
\emgr{Explain assumptions of analytic model but defer equations and derivations to Appendix \ref{sec:analytic}. Show numerical vs. analytic comparison plots and explain role of self gravity. }
\subsection{EOS Tables}
\emgr{Mention and briefly describe Saumon et al. EOS tables as future work. }

@@

\acknowledgements
\emgr{Acknowledge Ruth, Scott, Jonathan Fortney \& Didier Saumon.}  Portions of this project were supported by the {\it NASA} {\it Astrophysics Theory Program} and  {\it Origins of Solar Systems Program}  through grant NNX10AF35G. \emgr{JILA, Ruth's grants?}\\

\appendix
%\section{Extending the EOS tables}\label{sec:EOSext}
\section{Virial Equilibrium and Cooling}\label{sec:virial}
\section{Analytic Cooling Model}\label{sec:analytic}
\if\bibinc n
\bibliography{refs}
\fi

>>>>

\shorttitle{Critical Core Mass at Wide Separations}
\shortauthors{Youdin \& Piso \emph{TBD}}

\title{On the Minimum Core Mass for Giant Planet Formation}
%\title{Lower Limits on the Core Mass for Giant Planets}
%\title{The Critical Core Mass of Wide-Separation Giant Planets:  Lower Limits from Kelvin-Helmholtz Cooling}
\author{Ana-Maria Piso}
\affil{Harvard Smithsonian Center for Astrophysics}
\author{Andrew N.\ Youdin}
\affil{JILA, University of Colorado}
%\begin{abstract}
%\end{abstract}

\section{Introduction}
\emgr{Background topics to mention include: core accretion history, giant planets at wide separations and theoretical challenges posed (GI probably too big, core accretion maybe to slow).}

@@

\section{Atmosphere Models}\label{sec:model}

\subsection{Basic Assumptions}
\emgr{ Define and describe basic length scales; Describe disk conditions: MMSN etc. Describe basic assumptions: 1D atmosphere, smooth matching onto the disk, atmosphere composition etc.  \\
The idea is to start gentle with generalities and then get specific with the 2-layer quasistatic model after the basic eqns. have been given.}

%Describe assumptions specific to our model: quasistatic, $L_{KH} \gg L_{acc}$, two-layer, constant $L$. Explain how our model is intermediary between purely static models with $L=L_{acc}$ and full time-dependent evolutionary models.

\subsection{Structure Equations}
\emgr{Give in radius coordinates. EOS (polytrope, various assumption for $\nabla_{\mathrm{ad}}$ and $\mu$) and opacity laws (noting that we only consider cool radiative zones with dust opacities). Give boundary conditions}

\subsection{Standard Methods of Solution} 
\emgr{shooting vs. Henyey methods to satisfy above BC. Static models powered by accretion and full evolutionary calculations. Note that there's nothing really to explain about the ``shooting method" .  Fine to state the BC's and that we use the shooting method.  That's a difference between a paper and research notes. We probably want the radius version of structure ODEs.}

\subsection{Virial Equilibrium and Global Cooling}
\emgr{Start with basic idea of quasistatic contraction and $L = -\dot{E}$, then write out cooling equation and explain terms.   Explain where it comes from (local cooling + virial equilibrium).   Defer some derivation to Appendix \ref{sec:virial}.  In principle this is still general stuff, i.e. the global cooling eqn applies to all hydrostatic models, not just our quasi-static one.}

Before describing our quasistatic method we first describe the general ways that planetary Kelvin-Helmholz contraction differs from the stellar case.

\subsection{Quasi-static Two-layer Model}
\emgr{Two layer model and simplified treatment of radiative zone. How to obtain an evolutionary series from cooling model between subsequent static atmospheres. }

\section{Analytic Cooling Model}
\emgr{derive model (perhaps deferring some results to an appendix), give and explain expected scalings.  Could go after numerical results, but if first can use to explain some numerical results.}

\section{Quasistatic Kelvin-Helmholz Contraction}
\emgr{Show instantaneous profiles. Also L-t, L-M evolution.}

\subsection{The End of Quasistatic Evolution}
\emgr{Describe that time / luminosity becomes negative when $M_{atm} \sim M_c$.}

@@

\section{Neglected effects}
\emgr{EOS, time evolution, planetesimal accretion, inefficient convection (mixing length models).  Subsections as needed.}
\subsection{Planetesimal Accretion}
\emgr{Compare and contrast our results with Rafikov's, show accretion rates plot; discuss why our fixed core mass assumptions are valid in the regime we are considering.}

@@

\acknowledgements
\emgr{Acknowledge Ruth, Phil Armitage, Scott Kenyon, Jonathan Fortney \& Didier Saumon.}  Portions of this project were supported by the {\it NASA} {\it Astrophysics Theory Program} and  {\it Origins of Solar Systems Program}  through grant NNX10AF35G. \emgr{JILA, Ruth's grants?}\\

\appendix
%\section{Extending the EOS tables}\label{sec:EOSext}
\section{Derivation of Cooling Model}\label{sec:virial}
\section{Analytic Cooling Model Details}\label{sec:analytic}

\if\bibinc n
\bibliography{refs}
\fi
